### 2021-05-31 thesis/tex/oscillations.tex
update

%\input{commands}
%\begin{document}

\section{Neutrino Mixing}\label{ch:oscillation}
%TODO: small intro about PMNS and justification of oscillations
\subsection{Mass Generation}
As we saw in Eq.~\ref{eq:j_CC}
, the neutrino fields $\nu_\alpha$ only couple to the associated lepton fields $\ell_\alpha$, conserving the lepton number $L_\alpha$. We will now introduce two separate extensions to this part of the Standard Model.

We introduce a right-handed neutrino field, $\nu_R$. It has the usual properties of the conventional left-handed neutrino such as hypercharge and color zero. Moreover, since the electroweak gauge group $\text{SU}(2)_L$ only couple to left-handed particles and right-handed antiparticles, it transforms as a singlet under the SM symmetry group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. This neutrino is \emph{sterile} since it doesn't participate any of the SM interactions.

We extend the SM by adding a right-handed component to the Higgs-lepton Yukawa Lagrangian from Eq.~\ref{eq:YukawaLagrangian} with neutrino Yukawa couplings $Y_{\alpha \beta}^{\prime \nu}$ ,

\begin{align}
    \mathcal{L}_{H}=-\left( \frac{v + H}{\sqrt{2}} \right) \left[\ell_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime} + \nu_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \nu} \nu_{\beta R}^{\prime}\right]
\end{align}
Similar to how we diagonalized the lepton Yukawa couplings $Y_{\alpha \beta}^{\prime \ell}$ in Eq.~~\ref{eq:leptonYukawaDiag}, we diagonalize $Y_{\alpha \beta}^{\prime \nu}$ as
\begin{align}
    V_{\alpha k L}^{\nu \dagger} Y^{\prime \nu}_{\alpha \beta} V_{\beta j R}^{\nu}=Y^{\nu}_{kj} \,.
\end{align}
Now we introduce a crucial difference between the properties of the charged lepton and the neutrino fields.
While the charged lepton flavor eigenstate was uniquely determined by its mass eigenstate, the neutrino flavor is a superposition of mass eigenstates. This is because neutrinos are indirectly detected via the observation of its associated charged lepton, so there is no requirement of neutrino flavor eigenstates to have a definite mass. The flavor of a neutrino is then, by definition, the flavor of the associated charged lepton. This is commonly introduced as giving the mass eigenstates Latin numerals and letters, while the flavor eigenstates stay as Greek letters.

@@

Now, the Dirac neutrino field is
\begin{align}
    \nu_k = \nu_{kL} + \nu_{kR}\,.
\end{align}
Multiplying $\nu_k$ with its conjugate $\bar{\nu}_k$, we get 
\begin{align}
    \bar{\nu}_k \nu_k 
    & = \bar{\nu}_{k L} \nu_{k L} +\bar{\nu}_{k R}\nu_{k L} + \bar{\nu}_{k L}\nu_{k R} + \bar{\nu}_{k R}\nu_{k R} \nonumber \\
    & = \bar{\nu}_{k L}\nu_{k R} + \bar{\nu}_{k R}\nu_{k L} \nonumber \\
    & = \bar{\nu}_{k L}\nu_{k R} + \t{h.c.}
\end{align}
The same calculation for the charged lepton field yields the same result for $\ell_k$. Substituting this result and expanding the Higgs VEV into the fields gives us
\begin{align}
    \mathcal{L}_{H} 
    &=-\left( \frac{v + H}{\sqrt{2}} \right) \left[ y_{\alpha}^{\ell}   \bar{\ell}_\alpha \ell_\alpha  +  y_{k}^{\nu} \bar{\nu}_k \nu_k \right] \nonumber \\
    &=- \frac{y_{\alpha}^{\ell} v}{\sqrt{2}}   \bar{\ell}_\alpha \ell_\alpha   -  \frac{ y_{k}^{\nu} v}{\sqrt{2}} \bar{\nu}_k \nu_k  - \frac{y_{\alpha}^{\ell}}{\sqrt{2}}   \bar{\ell}_\alpha \ell_\alpha H  -  \frac{ y_{k}^{\nu}}{\sqrt{2}} \bar{\nu}_k \nu_k H\,.
\end{align}
Thus, this extension to the SM generates neutrino masses by the Higgs mechanism, in the same fashion as with the charged leptons and the quarks:
\begin{align}
    m_k = \frac{y_k^\nu v}{\sqrt{2}}
\end{align}

@@

Call $V^{\prime \nu \dagger}_{k \alpha}V^{\prime \ell}_{\alpha \alpha} = U^\dagger_{k \alpha}$. We will refer to the matrix $U$ built by the components $U_{k \alpha}$ as the
Pontecorvo-Maki-Nakagawa-Sakata (PMNS) matrix. %TODO: source
We now have 
\begin{align}\label{eq:j_CC3} %TODO: put sums on all above?
    j^\rho_L &= 2 \sum_\alpha \sum_k U^*_{\alpha k} \bar{\nu}_{k L} \gamma^\rho  \ell_{\alpha L}\,.
\end{align}
%TODO: add segue
\subsection{Neutrino Propagation}
Since we now know how the neutrino mass and flavor eigenstates combine, and have an expression for the flavor interaction with 
the neutrino's charged lepton partner, we are now ready to study the flavor oscillations themselves.

Now, the Fourier expansion of the field operator $\bar{\nu}_{k L}$ in the current of Eq.~\ref{eq:j_CC3} contains creation operators $a^\dagger_{\nu_k}$
of massive neutrinos with mass $m_k$. This means that the summation over the mass index $k$ constructs a flavor neutrino, which interacts with the charged lepton field $\ell_{\a L}$.
In other words, the charged current generates a flavor neutrino $\na$, which is a superposition of the mass eigenstates 
$\nu_k$ with weights $U_{\a k}^*$. In the ket-formalism, we express this as
\begin{align}\label{eq:osc_1}
    \ket{\na} = \sum_k U^*_{\a k} \ket{\nu_k}\,.
\end{align}
It is the mass eigenstates $\ket{\nu_k}$ that are eigenstates of the Hamiltonian, with eigenvalues
\begin{align}\label{eq:disp}
    E_k = \sqrt{\vec{p}^2 + m_k^2}\,.
\end{align}
The solution to the time-dependent Schrödinger equation 
\begin{align}\label{eq:TDSE}
    i \dv{}{t} \ket{\nu_k(t)} = H_0\ket{\nu_k(t)}\,,
\end{align}
where $H_0$ is the Hamiltonian and the subscript signifies that we are in vacuum.

The solution to Eq.~\ref{eq:TDSE} gives ut the time evolution
in the form of plane wave solutions: 
\begin{align}
    \ket{\nu_k(t)} = e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Inserting the plane wave solution into Eq.~\ref{eq:osc_1}, we get 
\begin{align}\label{eq:nat}
    \ket{\na(t)} = \sum_k U^*_{\a k} e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Now we know how to evolve and combine the mass eigenstates to form a flavor eigenstate, but how about the reverse?
 We swap the index $k\to j$ in Eq.~\ref{eq:osc_1} and multiply by $U_{\a k}$:
\begin{align}\label{eq:osc_k}
    \sum_\a U_{\a k}\ket{\na} &= \sum_{\alpha, j} U_{\a k} U_{\a j}^* \ket{\nu_j} \nonumber \\
    \sum_\a U_{\a k}\ket{\na} &= \sum_{j} \delta_{kj} \ket{\nu_j} \nonumber \\
    \sum_\a U_{\a k}\ket{\na} &= \ket{\nu_k}\,,
\end{align}
where we have used the unitarity of the leptonic mixing matrix. Eqs.~\ref{eq:osc_1} and \ref{eq:nat} yield 
\begin{align}
    \ket{\na(t)} &= \sum_k U^*_{\a k} e^{-iE_kt}\ket{\nu_k} \nonumber \\
    \ket{\na(t)} &= \sum_k U^*_{\a k} e^{-iE_kt}\left(\sum_\b U_{\b k}\ket{\nb}\right) \nonumber \\
    \ket{\na(t)} &= \sum_{k,\b} U^*_{\a k}U_{\b k} e^{-iE_kt} \ket{\nb}\,.
\end{align}
The probability of the flavor transition $\na \to \nb$ at time $t$ is $\abs{\bra{\nb}\ket{\na (t)}}^2$:
\begin{align}
    P_{\na \to \nb}(t) = \sum_{k,j} U^*_{\a k}U_{\b k}U^*_{\b j}U_{\a j} e^{-i(E_k-E_j)t}\,.
\end{align}

We assume the neutrino masses $m_k$ to be extremely small compared to their associated energies $E_k$. 
Thus, $v\sim 1$, and $\abs{\vec{p}} \sim E$ making the energy-dispersion relation of Eq.~\ref{eq:disp} to first order:
\begin{align}\label{eq:ultra_rel}
    E_k &= \sqrt{\vec{p}^2 + m_k^2} \nonumber \\
        &= \vec{p}^2\sqrt{1 + \frac{m_k^2}{\vec{p}^2}} \nonumber \\
        &\approx E + \frac{m_k^2}{2E}
\end{align}
Hence, the exponential can be simplified, and simplifying the notation $P_{\na \to \nb}(t) \to P_{\ab}(t)$ we get 
\begin{align}
    P_{\ab}(t) = \sum_{k,j} U^*_{\a k}U_{\b k}U^*_{\b j}U_{\a j} e^{-i(m_k^2-m_j^2)t/2E}\,.
\end{align}

Now, our approximation $v\approx 1$ implies $x\approx t $, thus
\begin{align}
    P_{\ab}(x) &= \sum_{k,j} U^*_{\a k}U_{\b k}U^*_{\b j}U_{\a j} e^{-i(m_k^2-m_j^2)x/2E} \nonumber \\
                       &= \sum_{k,j} U^*_{\a k}U_{\b k}U^*_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E})\,,
\end{align}
where we in the last step have defined the \emph{mass squared-difference} $\dm_{kj} = m_k^2 - m_j^2$. Since the oscillation probability depends on this quantity
rather than the individual masses, it is impossible
to measure the mass states $m_k$ through neutrino oscillations.
Squaring the unitarity condition $\sum_k U_{\a k}U_{\b k}^* = \delta_{\a \b}$ yields 
\begin{align}
    \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 = \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^* U_{\b k} U_{\a j} U_{\b j}^*]
\end{align}
In this work, we will always assume the mixing matrix $U$ to be real. 
We thus write 
\begin{align}\label{eq:Pab}%TODO: shorten?
    P_{\ab} &= \sum_{k,j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &=  \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 + \nonumber \\
            &\hspace{2em}+ \sum_{k \neq j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}U_{\a k} U_{\b k} U_{\b j} U_{\a j} \nonumber \\
            & \hspace{3em}+\sum_{k \neq j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}U_{\a k} U_{\b k} U_{\b j} U_{\a j} \nonumber \\
            &\hspace{3em}+ 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\exp(-i\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\cos(\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \sin^2{\left(\frac{\dm_{kj}x}{4E}\right)}\,,
\end{align}
which is the probability of neutrino vacuum oscillations. %TODO: maybe plot?
The calculation for antineutrinos yield the same result if one assumes the realness of the mixing matrix.

@@

We start with the effective Hamiltonian for the CC process. The Feynman rules for the left panel give us 
\begin{align}
    H^{CC} = \frac{G_F}{\sqrt{2}}\left[ \bar\ne \gamma^\rho\, (1- \gamma^5)\, e \right] \left[\bar e \,\gamma_\rho\, (1- \gamma^5)\, \ne \right]
\end{align}
By using the Fierz transformation 
\begin{align}
    \mathcal{L}^{V-A}(\psi_1,\psi_2,\psi_3,\psi_4) = \mathcal{L}^{V-A}(\psi_1,\psi_4,\psi_3,\psi_2)\,,
\end{align}
we can permute the terms inside the brackets, yielding
\begin{align}\label{eq:H_fierz}
    H^{CC} = \frac{G_F}{\sqrt{2}}\left[ \bar\ne \gamma^\rho\, (1- \gamma^5)\, \ne \right] \left[\bar e \,\gamma_\rho\, (1- \gamma^5)\, e \right]\,.
\end{align}
Now, lets consider a finite volume $V$ with electron states defined as 
\begin{align}\label{eq:e_states}
    \ket{e(p_e,h_e)} = \frac{1}{2E_eV}a_e^{(h_e)\dagger}(p_e)\ket{0}\,,
\end{align}
i.e. using the creation operator $a_e^{(h_e)\dagger}(p_e)$ to create electron states from vacuum with momenta $p_e$, energy $E_e$, and helicity $h_e$.
The density distribution of electrons in $V$ is $f(E_e,T)$, which we normalize to the total number of electrons as we integrate out the momenta $p_e$:
\begin{align}\label{eq:e_density}
    \int \dd p_e^3 f(E_e,T) = N_e V = n_e
\end{align}
Here, the electron density $N_e$ will ultimately determine the strength of the effective matter potential. 
To obtain the average effective Hamiltonian, project it on the electron states in Eq.~\ref{eq:e_states} and integrate over the density and sum over the helicities:

>>>>

% \input{commands}\begin{document}

\section{Neutrino Mixing}\label{ch:oscillation}
%TODO: small intro about PMNS and justification of oscillations
\subsection{Mass Generation}
We consider a right-handed neutrino field, $\nu_R$. It has the usual properties of the conventional left-handed neutrino such as hypercharge and color zero. Moreover, since the electroweak gauge group $\text{SU}(2)_L$ only couple to left-handed particles and right-handed antiparticles, $\nu_R$ transforms as a singlet under the SM symmetry group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. This neutrino field is \emph{sterile} since it doesn't participate any of the SM interactions.

We extend the Standard Model by adding a right-handed component of this field with neutrino Yukawa couplings $Y_{\alpha \beta}^{\prime \nu}$ to the Higgs-lepton Yukawa Lagrangian from Eq.~\ref{eq:YukawaLagrangian} 
\begin{align}
    \mathcal{L}_{H}=-\left( \frac{v + H}{\sqrt{2}} \right) \left[\ell_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime} + \nu_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \nu} \nu_{\beta R}^{\prime}\right]
\end{align}
Similar to how we diagonalized the lepton Yukawa couplings $Y_{\alpha \beta}^{\prime \ell}$ in Eq.~\ref{eq:leptonYukawaDiag}, we diagonalize $Y_{\alpha \beta}^{\prime \nu}$ as
\begin{align}
    V_{\alpha k L}^{\nu \dagger} Y^{\prime \nu}_{\alpha \beta} V_{\beta j R}^{\nu}=Y^{\nu}_{kj} \,.
\end{align}
Now we introduce a crucial difference between the properties of the charged lepton and the neutrino fields.
While the charged lepton flavor eigenstate was uniquely determined by its mass eigenstate, the neutrino flavor is a superposition of mass eigenstates. This is because neutrinos are indirectly detected via the observation of its associated charged lepton, so there is no requirement of neutrino flavor eigenstates to have a definite mass. The flavor of a neutrino is then, by definition, the flavor of the associated charged lepton. This is commonly introduced as giving the mass eigenstates Latin numerals and letters, while the flavor eigenstates stay as Greek letters.

@@

Now, the Dirac neutrino field is
\begin{align}
    \nu_k = \nu_{kL} + \nu_{kR}\,.
\end{align}
Multiplying $\nu_k$ with its conjugate $\bar{\nu}_k$, we get 
\begin{align}
    \bar{\nu}_k \nu_k 
    & = \bar{\nu}_{k L} \nu_{k L} +\bar{\nu}_{k R}\nu_{k L} + \bar{\nu}_{k L}\nu_{k R} + \bar{\nu}_{k R}\nu_{k R} \nonumber \\
    & = \bar{\nu}_{k L}\nu_{k R} + \bar{\nu}_{k R}\nu_{k L} \nonumber \\
    & = \bar{\nu}_{k L}\nu_{k R} + \t{h.c.}
\end{align}
The same calculation for the charged lepton field yields the same result for $\ell_k$. Substituting this result and expanding the Higgs vacuum expectation value $v$ into the fields gives us
\begin{align}
    \mathcal{L}_{H} 
    &=-\left( \frac{v + H}{\sqrt{2}} \right) \left[ y_{\alpha}^{\ell}   \bar{\ell}_\alpha \ell_\alpha  +  y_{k}^{\nu} \bar{\nu}_k \nu_k \right] \nonumber \\
    &=- \frac{y_{\alpha}^{\ell} v}{\sqrt{2}}   \bar{\ell}_\alpha \ell_\alpha   -  \frac{ y_{k}^{\nu} v}{\sqrt{2}} \bar{\nu}_k \nu_k  - \frac{y_{\alpha}^{\ell}}{\sqrt{2}}   \bar{\ell}_\alpha \ell_\alpha H  -  \frac{ y_{k}^{\nu}}{\sqrt{2}} \bar{\nu}_k \nu_k H\,.
\end{align}
Thus, this extension to the SM generates neutrino masses by the Higgs mechanism, in the same fashion as with the charged leptons and the quarks:
\begin{align}
    m_k = \frac{y_k^\nu v}{\sqrt{2}}
\end{align}

@@

Call $V^{\prime \nu \dagger}_{k \alpha}V^{\prime \ell}_{\alpha \alpha} = U^\dagger_{k \alpha}$. We will refer to the matrix $U$ built by the components $U_{k \alpha}$ as the
Pontecorvo-Maki-Nakagawa-Sakata (PMNS) matrix. %TODO: source
We now have 
\begin{align}\label{eq:j_CC3} %TODO: put sums on all above?
    j^\rho_L &= 2 \sum_\alpha \sum_k U^\dagger_{\alpha k} \bar{\nu}_{k L} \gamma^\rho  \ell_{\alpha L}\,.
\end{align}

What form does this new matrix $U$ take? By construction, it provides the unitary transformation between the flavor and mass bases.
Any unitary $3\times3$ matrix can be parametrized using three mixing angles and six phases. However, not all phases affect the 
charged and weak currents and are thus not observable. Moreover, both the Lagrangian and the currents are invariant under global $U(1)$ transformations,
leaving only one physical phase for us to measure. Thus we are down to four degrees of freedom in the three dimensional case: three mixing angles
of the form $\sin{(\theta_{ij})}/\cos{(\theta_{ij})} = s_{ij}/c_{ij}$, and one phase of the form $e^{i\delta}$.

We construct the PMNS matrix using the rotation matrixes from $SO(3)$ as
\begin{align}
    U &= R_{23}R_{13,\delta_{\text{CP}}}R_{12} \nonumber \\
      & = 
    \begin{pmatrix}{ccc}1 & 0 & 0 \\ 0 & c_{23} & s_{23} \\ 0 & -s_{23} & c_{23}\end{pmatrix}
\begin{pmatrix}{ccc}c_{13} & 0 & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 0 & 1 & 0 \\ -s_{13} e^{i \delta_{\mathrm{CP}}} & 0 & c_{13}\end{pmatrix}
\begin{pmatrix}{ccc}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1\end{pmatrix}\,.
\end{align}

The term $\delta_\text{CP}$ determines the degree of charge-parity violation. In this work, we will assume CP invariance by always
setting $\delta_{CP} = \SI{0}{\degree}$. Thus,

\begin{align}
    U &=
    \begin{pmatrix}1 & 0 & 0 \\ 0 & c_{23} & s_{23} \\ 0 & -s_{23} & c_{23}\end{pmatrix}
\begin{pmatrix}c_{13} & 0 & s_{13} \\ 0 & 1 & 0 \\ -s_{13}  & 0 & c_{13}\end{pmatrix}
\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1\end{pmatrix} \nonumber \\
    &= 
    \begin{pmatrix}c_{12} c_{13} & s_{12} c_{13} & s_{13} \\ 
        -s_{12} c_{23}-c_{12} s_{23} s_{13} & c_{12} c_{23}-s_{12} s_{23} s_{13} & s_{23} c_{13} \\ 
        s_{12} s_{23}-c_{12} c_{23} s_{13}  & -c_{12} s_{23}-s_{12} c_{23} s_{13} & c_{23} c_{13}
    \end{pmatrix}\,.
\end{align}

In this work, we will be using the best-fit values from~\cite{nufit} with the exception of the CP-violating phase $\delta_\text{CP}$. The $90\% $ CL values are
\begin{align}\label{eq:nufitparams}
    \theta_{12} = \SI{33.44}{\degree},\hspace{1em} \theta_{13} = \SI{8.57}{\degree},\hspace{1em} \theta_{23} = \SI{49.2}{\degree}, \hspace{1em} \delta_\text{CP} = 0\,.
\end{align}
The PMNS matrix is then real, with numerical values 
\begin{align}\label{eq:Uvalues}
    U = \begin{pmatrix}
        U_{e 1} & U_{e2} & U_{e3} \\
        U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
        U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
    \end{pmatrix} 
    = \begin{pmatrix}
        0.825 & 0.545 & 0.149 \\
        -0.455 & 0.485 & 0.746 \\
        0.334 & -0.684 & 0.649
    \end{pmatrix} \,.
\end{align}
In our study, we will only use atmospheric neutrinos, of which $\nm$ are the most abundant. We see that $U_{\mu\tau}$ is the largest $U_{\mu\beta}$ term,
suggesting that $\nm$ primarily mixes into $\nt$, at least in vacuum. It turns out that matter effects indeed preserves this ordering, making the $\nm \to \nt$ transition
the most abundant. This means we will be able to stringently constrain parameters relating to $\mu\tau$ mixing.

\subsection{Neutrino Propagation}
Since we now know how the neutrino mass and flavor eigenstates combine, and have an expression for the flavor interaction with 
the neutrino's charged lepton partner, we are now ready to study the flavor oscillations themselves.

Now, the Fourier expansion of the field operator $\bar{\nu}_{k L}$ in the current of Eq.~\ref{eq:j_CC3} contains creation operators $a^\dagger_{\nu_k}$
of massive neutrinos with mass $m_k$. This means that the summation over the mass index $k$ constructs a flavor neutrino, which interacts with the charged lepton field $\ell_{\a L}$.
In other words, the charged current generates a flavor neutrino $\na$, which is a superposition of the mass eigenstates 
$\nu_k$ with weights $U_{\a k}^\dagger$. In the ket-formalism, we express this as
\begin{align}\label{eq:osc_1}
    \ket{\na} = \sum_k U^\dagger_{\a k} \ket{\nu_k}\,.
\end{align}
It is the mass eigenstates $\ket{\nu_k}$ that are eigenstates of the Hamiltonian, with eigenvalues
\begin{align}\label{eq:disp}
    E_k = \sqrt{\vec{p}^2 + m_k^2}\,.
\end{align}
The solution to the time-dependent Schrödinger equation 
\begin{align}\label{eq:TDSE}
    i \dv{}{t} \ket{\nu_k(t)} = H_0\ket{\nu_k(t)}\,,
\end{align}
where $H_0$ is the Hamiltonian and the subscript signifies that we are in vacuum.

The solution to Eq.~\ref{eq:TDSE} gives ut the time evolution
in the form of plane wave solutions: 
\begin{align}
    \ket{\nu_k(t)} = e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Inserting the plane wave solution into Eq.~\ref{eq:osc_1}, we get 
\begin{align}\label{eq:nat}
    \ket{\na(t)} = \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k}\,.
\end{align}
Now we know how to evolve and combine the mass eigenstates to form a flavor eigenstate, but how about the reverse?
 We swap the index $k\to j$ in Eq.~\ref{eq:osc_1} and multiply by $U_{\a k}$:
\begin{align}\label{eq:osc_k}
    \sum_\a U_{\a k}\ket{\na} &= \sum_{\alpha, j} U_{\a k} U_{\a j}^\dagger \ket{\nu_j} \nonumber \\
    \sum_\a U_{\a k}\ket{\na} &= \sum_{j} \delta_{kj} \ket{\nu_j} \nonumber \\
    \sum_\a U_{\a k}\ket{\na} &= \ket{\nu_k}\,,
\end{align}
where we have used the unitarity of the leptonic mixing matrix. Eqs.~\ref{eq:osc_1} and \ref{eq:nat} yield 
\begin{align}
    \ket{\na(t)} &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\ket{\nu_k} \nonumber \\
    \ket{\na(t)} &= \sum_k U^\dagger_{\a k} e^{-iE_kt}\left(\sum_\b U_{\b k}\ket{\nb}\right) \nonumber \\
    \ket{\na(t)} &= \sum_{k,\b} U^\dagger_{\a k}U_{\b k} e^{-iE_kt} \ket{\nb}\,.
\end{align}
The probability of the flavor transition $\na \to \nb$ at time $t$ is $\abs{\bra{\nb}\ket{\na (t)}}^2$:
\begin{align}
    P_{\na \to \nb}(t) = \sum_{k,j} U^\dagger_{\a k}U_{\b k}U^*_{\b j}U_{\a j} e^{-i(E_k-E_j)t}\,.
\end{align}

We assume the neutrino masses $m_k$ to be extremely small compared to their associated energies $E_k$. 
Thus, $v\sim 1$, and $\abs{\vec{p}} \sim E$ making the energy-dispersion relation of Eq.~\ref{eq:disp} to first order:
\begin{align}\label{eq:ultra_rel}
    E_k &= \sqrt{\vec{p}^2 + m_k^2} \nonumber \\
        &= \vec{p}^2\sqrt{1 + \frac{m_k^2}{\vec{p}^2}} \nonumber \\
        &\approx E + \frac{m_k^2}{2E}
\end{align}
Hence, the exponential can be simplified, and simplifying the notation $P_{\na \to \nb}(t) \to P_{\ab}(t)$ we get 
\begin{align}
    P_{\ab}(t) = \sum_{k,j} U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} e^{-i(m_k^2-m_j^2)t/2E}\,.
\end{align}

Now, our approximation $v\approx 1$ implies $x\approx t $, thus
\begin{align}
    P_{\ab}(x) &= \sum_{k,j} U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} e^{-i(m_k^2-m_j^2)x/2E} \nonumber \\
                       &= \sum_{k,j} U^\dagger_{\a k}U_{\b k}U^\dagger_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E})\,,
\end{align}
where we in the last step have defined the \emph{mass squared-difference} $\dm_{kj} = m_k^2 - m_j^2$. Since the oscillation probability depends on this quantity
rather than the individual masses, it is impossible
to measure the mass $m_k$ through neutrino oscillations.
Squaring the unitarity condition $\sum_k U_{\a k}U_{\b k}^\dagger = \delta_{\a \b}$ yields 
\begin{align}
    \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 = \delta_{\a \b} - 2 \sum_{k>j}\Re[U_{\a k}^\dagger U_{\b k} U_{\a j} U_{\b j}^\dagger]
\end{align}

Since we take the PMNS matrix to be real and thus Hermitian, we drop the complex transpose in the following calculation. We have
\begin{align}\label{eq:Pab}%TODO: shorten?
    P_{\ab} &= \sum_{k,j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &=  \sum_k \abs{U_{\a k}}^2\abs{U_{\b k}}^2 + \nonumber \\
            &\hspace{2em}+ \sum_{k \neq j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}U_{\a k} U_{\b k} U_{\b j} U_{\a j} \nonumber \\
            & \hspace{3em}+\sum_{k \neq j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2 \sum_{k>j}U_{\a k} U_{\b k} U_{\b j} U_{\a j} \nonumber \\
            &\hspace{3em}+ 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \exp(-i\frac{\dm_{kj}x}{2E}) \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\exp(-i\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \left[1-\cos(\frac{\dm_{kj}x}{2E})\right] \nonumber \\
            &= \delta_{\a \b} - 2\sum_{k > j} U_{\a k}U_{\b k}U_{\b j}U_{\a j} \sin^2{\left(\frac{\dm_{kj}x}{4E}\right)}\,,
\end{align}
which is the probability of neutrino vacuum oscillations. %TODO: maybe plot?
The calculation for antineutrinos yield the same result if one continues to assume the Hermicity of the mixing matrix.

For our purposes, the probabilities in Eq.~\ref{eq:Pab} are rather uninteresting, since they are only valid in vacuum.
Despite this fact, this form elucidates an important aspect of neutrino oscillations. The mixing matrix elements
determine the amplitude of the oscillations, while the mass-squared differences together with the ratio $L/E$ determine the 
frequency.

@@

We start with the effective Hamiltonian for the CC process. The Feynman rules for the left panel give us 
\begin{align}
    H_{CC} = \frac{G_F}{\sqrt{2}}\left[ \bar\ne \gamma^\rho\, (1- \gamma^5)\, e \right] \left[\bar e \,\gamma_\rho\, (1- \gamma^5)\, \ne \right]
\end{align}
By using the Fierz transformation 
\begin{align}
    \mathcal{L}^{V-A}(\psi_1,\psi_2,\psi_3,\psi_4) = \mathcal{L}^{V-A}(\psi_1,\psi_4,\psi_3,\psi_2)\,,
\end{align}
we can permute the terms inside the brackets, yielding
\begin{align}\label{eq:H_fierz}
    H_{CC} = \frac{G_F}{\sqrt{2}}\left[ \bar\ne \gamma^\rho\, (1- \gamma^5)\, \ne \right] \left[\bar e \,\gamma_\rho\, (1- \gamma^5)\, e \right]\,.
\end{align}
Now, lets consider a finite volume $V$ with electron states defined as 
\begin{align}\label{eq:e_states}
    \ket{e(p_e,h_e)} = \frac{1}{2E_eV}a_e^{(h_e)\dagger}(p_e)\ket{0}\,,
\end{align}
i.e. using the creation operator $a_e^{(h_e)\dagger}(p_e)$ to create electron states from vacuum with momenta $p_e$, energy $E_e$, and helicity $h_e$.
The density distribution of electrons in $V$ is $f(E_e,T)$, which we normalize to the total number of electrons as we integrate out the momenta $p_e$:
\begin{align}\label{eq:e_density}
    \int \dd p_e^3 f(E_e,T) = N_e V = n_e
\end{align}
Here, the electron density $N_e$ will ultimately determine the strength of the effective matter potential. 
To obtain the average effective Hamiltonian, project it on the electron states in Eq.~\ref{eq:e_states} and integrate over the density and sum over the helicities:
